---
title: "ACCEPT 3.0-CPRD: Predicting COPD Exacerbation Risk in UK Primary Care"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ACCEPT 3.0-CPRD: Predicting COPD Exacerbation Risk in UK Primary Care}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(accept)
```

## Overview

`accept3_cprd()` implements ACCEPT 3.0-CPRD, a recalibration of the ACCEPT 2.0
model specifically for UK primary-care patients. It was derived from the
**Clinical Practice Research Datalink (CPRD)**, a large UK primary-care database,
and is designed to produce well-calibrated 1-year exacerbation risk predictions
in this setting.

The function predicts two outcomes:

- **Moderate-to-severe exacerbation** probability and rate
- **Severe exacerbation** probability and rate

It can also be called via the unified `accept()` interface using
`country = "GBR-primary"`.

---

## When to Use ACCEPT 3.0-CPRD

| Model | Population | Setting |
|-------|-----------|---------|
| `accept2()` | Canadian/international | Hospital or specialist |
| `accept3()` / `accept3_gbr()` | UK | Secondary care (spirometry clinic) |
| **`accept3_cprd()`** | **UK** | **Primary care (GP)** |

Use `accept3_cprd()` when your patients are from a **UK general practice** setting,
where disease severity tends to be milder and fewer specialist measurements are
routinely available.

---

## Required Identifier

The following column is required but is **not** a predictor — it is used solely
to label rows in the output so predictions can be matched back to patients:

| Column | Type | Description |
|--------|------|-------------|
| `ID` | character/numeric | Unique patient identifier |

---

## Required Predictors

The following columns are the actual **model inputs** and must be present in your data:

| Column | Type | Description |
|--------|------|-------------|
| `age` | numeric | Age in years |
| `male` | integer (0/1) | Sex (1 = male) |
| `FEV1` | numeric | FEV1 % predicted |
| `LastYrExacCount` | integer | Total exacerbations in the past year (moderate + severe) |
| `LastYrSevExacCount` | integer | Severe (hospitalised) exacerbations in the past year |
| `mMRC` **or** `SGRQ` | numeric | Dyspnoea score (mMRC 0–4) or SGRQ total score |

> **Note:** `LastYrSevExacCount` must not exceed `LastYrExacCount`. At least one
> of `mMRC` or `SGRQ` must be provided; if only `SGRQ` is available it is
> converted to `mMRC` internally.

---

## Optional Predictors

These columns are used if available; if missing they are **automatically imputed**
using a CPRD-specific sequential triangular regression model:

| Column | Type | Description |
|--------|------|-------------|
| `LABA` | integer (0/1) | Long-acting beta-agonist use |
| `oxygen` | integer (0/1) | Long-term oxygen therapy |
| `ICS` | integer (0/1) | Inhaled corticosteroid use |
| `LAMA` | integer (0/1) | Long-acting muscarinic antagonist use |
| `statin` | integer (0/1) | Statin or CVD medication use |
| `BMI` | numeric | Body mass index (clamped to 10–60) |
| `smoker` | integer (0/1) | Current smoker |

Imputation proceeds in the order above. Each variable is predicted from the
mandatory predictors plus any previously imputed optional predictors.

---

## Basic Usage

```{r basic, eval = FALSE}
library(accept)

# Using the sample patients bundled with the package
results <- accept3_cprd(samplePatients)
head(results)
```

```{r basic_show, echo = FALSE}
# Example output structure (not run)
cat(
"# A tibble: 6 x 4
  predicted_exac_probability predicted_exac_rate predicted_severe_exac_probability predicted_severe_exac_rate
                       <dbl>               <dbl>                             <dbl>                      <dbl>
1                      0.699               1.20                              0.356                      0.439
2                      0.324               0.392                             0.0415                     0.0424
...")
```

---

## Accessing Results via `accept()`

You can also call `accept3_cprd` through the unified interface:

```{r unified, eval = FALSE}
results <- accept(
  newdata  = samplePatients,
  version  = "accept3",
  country  = "GBR-primary"
)
```

---

## Including Predictors in the Output

Set `return_predictors = TRUE` to include the input columns alongside predictions.
This is useful for checking which values were imputed:

```{r return_preds, eval = FALSE}
results_full <- accept3_cprd(samplePatients, return_predictors = TRUE)
names(results_full)
```

---

## Imputation Messages

By default, imputation is silent. Set `quiet = FALSE` to see which predictors
were imputed for each patient:

```{r quiet, eval = FALSE}
results <- accept3_cprd(samplePatients, quiet = FALSE)
# Example message:
# Patient 3: imputed LABA, BMI, smoker
```

---

## Handling Missing mMRC / SGRQ

The function requires either `mMRC` (0–4) or `SGRQ` (0–100). If only `SGRQ`
is provided, it is converted to `mMRC` using the published conversion formula
before prediction. If neither is available, an error is raised.

```{r sgrq, eval = FALSE}
# Data with SGRQ instead of mMRC
patients_sgrq <- samplePatients
patients_sgrq$mMRC <- NULL  # remove mMRC
patients_sgrq$SGRQ <- c(45, 30, 60, 20, 55, 10)  # add SGRQ

results <- accept3_cprd(patients_sgrq)
```

---

## Recalibration Method

ACCEPT 3.0-CPRD applies a Cox-model recalibration to ACCEPT 2.0 predictions:

$$\hat{p}_{UK} = 1 - \exp\!\left(-H_0 \cdot \exp\!\left(\beta \cdot \log(-\log(1-\hat{p}_{2}))\right)\right)$$

where $\hat{p}_{2}$ is the ACCEPT 2.0 predicted probability, and $H_0$ and
$\beta$ are optimism-corrected parameters estimated via 200-resample bootstrap
from the CPRD dataset:

| Outcome | $H_0$ | $\beta$ |
|---------|-------|---------|
| Moderate-to-severe | 0.676 | 0.986 |
| Severe | 0.482 | 1.124 |

---

## Comparing Models

```{r compare, eval = FALSE}
library(accept)
library(dplyr)

patients <- samplePatients

a2   <- accept2(patients)
a3   <- accept3_cprd(patients)

comparison <- bind_cols(
  ID        = patients$ID,
  accept2   = a2$predicted_exac_probability,
  accept3uk = a3$predicted_exac_probability
)

print(comparison)
```

Predictions from ACCEPT 3.0-CPRD are generally **lower** than ACCEPT 2.0 for
moderate-to-severe exacerbations, reflecting the milder disease burden typically
seen in UK primary care compared with specialist settings.

---

## See Also

- `accept()` — unified interface for all ACCEPT model versions
- `accept3()` / `accept3_gbr()` — ACCEPT 3.0 UK secondary-care version
- `samplePatients` — example dataset


---

## Session Info

```{r session, eval = FALSE}
sessionInfo()
```
